Unifying quasars with clumpy wind models

### 2015-08-28 paper_mnras.tex
improved figure captions

The spectra of 
quasars and luminous active galactic nuclei (AGN) 
typically exhibit a series of strong emission lines
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is often attributed to emssion from a 
geometrically thin, optically thick accretion disc surrounding the central black hole, similar
to that described by \cite{shakurasunyaev1973}.
In addition to the inflowing accreting material, i
outflows are ubiquitous in AGN
and quasars \citep{kellerman1989,ganguly2008}. These outflows can take the form of 
highly collimated radio jets \citep[e.g.][]{hazard1963,potash1980,perley1984,marscher2006}, 
or mass-loaded `winds' emanating from the accretion disc 
\citep{weymann1991,turnermiller2009}. 
Outflows in AGN offer a 
potential feedback mechanism through which the central source can 
affect its environment \citep{king2003,king2005,fabian2012}
-- feedback that is required in models of galaxy evolution \citep{springel2005}
and may explain the `$M-\sigma$' relation \citep{silkrees1998,haring2004}.

@@

To treat line transfer, we adopt the same hybrid scheme 
described by M15, 
in which the energy flows
through the system are described in terms of indivisible
energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section 2.3).
These energy packets interact with either two-level `simple ions'
or full `macro-atoms'. The macro-atom implementation 
is described in full by \cite{lucy2002, lucy2003}.
Our scheme allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements which are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions (see M15). In this study,
only H \& He are treated as macro-atoms, because 
we expect recombination to be important
in determining their level populations and resultant line emission.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/uvspec.png}
\caption
{
Synthetic spectra at five viewing angles in our. The coloured lines
show different quasar and BAL quasar composites, and the dotted line shows a disc
only continuum to show the effect of the outflow on the continuum level.
}
\label{fig:uvspec}
\end{figure*} %fullpage

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/windnew3.png}
\caption
{
A cartoon describing the broad classes of sightline 
in our model, illustrates how geometric effects lead to 
the different emergent spectra. The colour gradient is approximate,
but indicates the stratified ionization structure, 
from highly ionized (yellow) to low ionization (purple) material.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:sightline}
\end{figure}

@@

When comparing BALQSO and quasar composites, it is apparent
that they possess remarkably similar line strengths and widths 
\citep[e.g.][]{reichard2003}.
This presents a challenge to our model, as well as the geometric 
unification picture in general.
Limb darkening and foreshortening causes the disc continuum emission to be 
strongly anisotropic, yet the line emission in our models is much more isotropic. 
This has the effect of enhancing the line-to-continuum ratios at high inclinations. 
To construct a scenario where emission 
line equivalent widths are comparable at all inclinations requires 
significant fine-tuning when considering a classical thin disc. 
An alternative, simpler solution is that the emergent continuum is roughly isotropic.

General relativistic effects -- specifically, light bending
and relativistic beaming -- can cause 
the accretion disc SED to become more isotropic \citep[e.g.][]{zhang1997,munozdarias2013}.
To generate GR disc spectra, we use the code \agn\ \citep{hubeny2000,davishubeny2006,davis2007}. 
The output flux at three different wavelengths
as a function of inclination for an \agn\ model with the same disc and BH parameters
as our clumpy wind model is shown in figure~\ref{fig:f2000}.
The effects of GR on an AGN disc are much less extreme 
in the UV portion of the spectrum than the calculations
by \cite{zhang1997} for the X-rays in X-ray binaries.
As a result, the disc emission is still strongly anisotropic.
GR alone therefore cannot explain the line ratio trends in quasars by making the disc
emit more isotropically.
This supports the findings of \cite{risaliti2011}, who find that EW distributions 
in quasars are consistent with anisotropic emission from optically thick, 
disc-like structures for {\em both} the continuum source and BLR. 
We will explore these ideas further in a future study.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/lx.png}
\caption
{
X-ray properties of the H13 (purple squares) and clumped model (red
squares), plotted against monochromatic luminosity 
at 2500\AA. The points are labeled according to inclination; angles
$>70^\circ$ correspond to BALs in our scheme.
Also plotted are the samples considered by Saez et al. 2012 on a similar plot; 
The COMBI-7 AGN and the BQS samples Steffen et al. (2006) and the Saez et al. (2012) 
sample of BALQSOs. The dotted lines show the best fit relations for non-BALQSOs 
from Steffen et al. (2006).
}
\label{fig:xray}
\end{figure*} %fullpage

@@

% These line profiles also indicate a potential problem with our model. 
% \cite{odowd2015} find that the LoBAL profiles in H1413+117 have a 
% higher velocity onset than the higher ionization BALs, suggesting that the
% wind becomes less ionized as radius increases. 
% We find the opposite trend in velocity onsets. In addition, examination of 
% figure~\ref{fig:wind} shows that ionization parameter decreases with radius, 
% because the decrease in density wins over geometric dilution of the radiation 
% field. While the \cite{odowd2015} hypothesis comes from one object,
% it is clear that {\em at least some} BALQSOs must become less ionized at 
% higher velocities, contrary to what one expects from our modelling.

% DISCUSSION

% This could potentially be achieved by wind reprocessing or a more realistic, GR treatment
% of the accretion disc. We briefly discuss both these possibilities.

% \subsection{Wind reprocessing}

% If the majority of continuum radiation was reprocessed by nearby photoionized plasma (such 
% as the outflow itself), then this would result in a roughly
% isotropic radiation field. There are a number of appealing aspects to this scenario.
% First, it may go some way to explaining recent microlensing (REFs) and timing (REFs)
% observations which suggest that accretion disc sizes are a factor of $\sim3$ 
% larger than one expects from a Shakura \& Sunyaev (1973) thin disc. 
% Second, it implies a wide angle, large covering factor wind in order to 
% intercept most of the disc continuum. This would make the outflow 
% an effective feedback mechanism \citep[e.g.][]{borguet2012}, 
% and would imply that BALQSO outflows
% have a similar geometry to the ultra-fast outflows seen in nearby AGN 
% \citep{reeves2003,poundsreeves2009,tombesi2010a}.
% We note that spectropolarimetry of BALQSOs shows that the light from BAL troughs is $\sim10\%$ 
% linearly polarised, but the continuum generally has negligible polarisation \citep{lamy2000}.
% This suggests that the continuum of BALQSOs cannot be significantly enhanced by
% {\em single} scattering into the line of sight, but does not rule out 
% {\em multiple} scatterings leading to a more isotropic continuum.

% \subsection{General relativistic effects}

% General relativistic effects -- specifically, light bending
% and relativistic beaming -- can cause 
% the accretion disc SED to become more isotropic \citep[e.g.][]{zhang1997,munozdarias2013}.
% To generate GR disc spectra, we use the code \agn\ \citep{hubeny2000,davishubeny2006,davis2007}. 
% The output flux at three different wavelengths
% as a function of inclination for an \agn\ model with the same disc and BH parameters
% as our clumpy wind model is shown in figure~\ref{fig:f2000}.
% The effects of GR on an AGN disc are much less extreme 
% in the UV portion of the spectrum than the calculations
% by \cite{zhang1997} for the X-rays in X-ray binaries.
% As a result, the disc emission is still strongly anisotropic.
% GR alone therefore cannot explain the line ratio trends in quasars.

% To examine the effect of GR on our models, we carry out the following procedure.
% First, we fit our output spectrum with a polynomial. This continuum fit can 
% be subtracted from the actual spectrum 
% to produce a model of the line emission only. We then calculate the 
% `wind continuum factor', by comparing the model fit with the emergent 
% disc spectrum without a wind. This factor gives us an approximate way to quantify the 
% effect of the wind on a disc continuum. It will generally be $>1$ at high inclinations,
% and $<1$ at low inclinations. This calculates emergent accretion disc spectra from an AGN disc atmosphere
% including relativistic effects. Finally, produce our final synthetic spectrum 
% by convolving this the line emission spectrum with the \agn\ model for the model parameters.

% Figure 10 shows the resultant spectrum. Including GR clearly helps

% Can GR effects offer a potential solution? (not quite!)

% {\bf Figures 8 and 9: AGNSPEC F(2000) as a function of viewing angle compared to a BB disc.
% spectrum compared to composites with AGNSPEC correction.}

% \begin{figure*} %fullpage
% \centering
% \includegraphics[width=1.0\textwidth]{figures/adjusted_withbals.png}
% \caption
% {
% $F_\lambda$ normalised to $F_{2000}$. {\bf Again, this is a placeholder- but I'm thinking some kind of comparison to composite including the adjusted continuum, showing that we can't get it exactly right}.
% }
% \label{fig:finalcomp}
% \end{figure*} %fullpage

% \subsection{The BALQSO fraction and wind covering factor}
% \label{sec:balfrac}

% The disc SED has a profound effect on the intrinsic BAL fraction
% inferred from flux-limited samples. This in turn impacts on 
% the geometry and covering factor assumed for BALQSO and unified wind models,
% and even effects the efficiency of feedback.
% \cite{krolik1998} noted that the effect of limb darkening and foreshortening
% would result in a signicantly underestimated covering factor for BALQSO winds.
% The effect is complicated further by the quasar luminosity function -- if 
% BALQSOs have lower observed luminosities for given accretion disc properties 
% then they may be drawn from a different bin in the quasar luminosity function 
% \citep{goodrich1997}. This will also have a similar impact
% on the LoBALQSOs and FeLoBALQSOs sub-populations.
% The complex (and as yet unknown) selection effects are understandably 
% not taken into account by current estimates of the intrinsic BALQSO fraction 
% \citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.

% The way forward is somewhat unclear. Given the lack of available observational 
% constraints, models cannot readily be used to inform the selection criteria when 
% calculating the BALQSO fraction. There are two main possibilities:
% First, the disc emission is roughly isotropic, in which case the BAL fraction
% estimates should be a good estimate of the covering factor of the outflow,
% but the question about how this isotropy is achieved remains outstanding.
% Or second, the disc emission is strongly anisotropic, which follows more
% naturally from a disc geometry but has large implications
% for the inferred covering factors, as well as leaving the similar line strengths
% in BAL and non-BAL quasar compoosites unexplained (see previous sections).
% Understanding the true nature of the disc SED, including the
% angular distribution of radiation, is clearly crucial in order
% to accurately estimate the true covering factor of BALQSO winds.

% A brief comment, citing Goodrich / Krolik \& Voit on the 
% way in which anisotropy / attenuation affect the inferred
% BAL fraction. We also need to be aware that there will be a number of selection
% effects in building up the composites, and we should discuss these
% and the subtleties involved.

% \begin{figure}
% \centering
% \includegraphics[width=0.45\textwidth]{figures/f2000_m9.png}
% \caption
% {
% $F_\lambda$ at $2000$~\AA\ as a function of inclination from
% \agn\ models. Spectra are computed for BH of $10^9~M_\odot$ with
% a number of different BH spins and Eddington fractions. The black line
% shows a standard multi-temperature blackbody AD model.
% }
% \label{fig:alpha_ox}
% \end{figure}

>>>>

The spectra of 
quasars and luminous active galactic nuclei (AGN) 
typically exhibit a series of strong emission lines
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is often attributed to emssion from a 
geometrically thin, optically thick accretion disc surrounding the central black hole, similar
to that described by \cite{shakurasunyaev1973}.
In addition to the inflowing accreting material,
outflows are ubiquitous in AGN
and quasars \citep{kellerman1989,ganguly2008}. These outflows can take the form of 
highly collimated radio jets \citep[e.g.][]{hazard1963,potash1980,perley1984,marscher2006}, 
or mass-loaded `winds' emanating from the accretion disc 
\citep{weymann1991,turnermiller2009}. 
Outflows in AGN offer a 
potential feedback mechanism through which the central source can 
affect its environment \citep{king2003,king2005,fabian2012}
-- feedback that is required in models of galaxy evolution \citep{springel2005}
and may explain the `$M-\sigma$' relation \citep{silkrees1998,haring2004}.

@@

To treat line transfer, we adopt the same hybrid scheme 
described by M15, 
in which the energy flows
through the system are described in terms of indivisible
energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section 2.3).
These energy packets interact with either two-level `simple ions'
or full `macro-atoms'. The macro-atom implementation 
is described in full by \cite{lucy2002, lucy2003}.
Our scheme allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements which are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions (see M15). In this study,
only H is treated as a macro-atom, because 
we expect recombination to be important
in determining its level populations and resultant line emission.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/uvspec.png}
\caption
{
Synthetic spectra at five viewing angles in our model. The coloured lines
show different quasar and BAL quasar composites, and the dotted line shows a disc
only continuum to show the effect of the outflow on the continuum level.
}
\label{fig:uvspec}
\end{figure*} %fullpage

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/windnew3.png}
\caption
{
A cartoon describing the broad classes of sightline 
in our model, illustrating how geometric effects lead to 
the different emergent spectra. The colour gradient is approximate,
but indicates the stratified ionization structure, 
from highly ionized (yellow) to low ionization (purple) material.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:sightline}
\end{figure}

@@

When comparing BALQSO and quasar composites, it is apparent
that they possess remarkably similar line strengths and widths 
\citep[e.g.][]{weymann1991,reichard2003}.
This presents a challenge to our model, as well as the geometric 
unification picture in general.
Limb darkening and foreshortening causes the disc continuum emission to be 
strongly anisotropic, yet the line emission in our models is much more isotropic. 
This has the effect of enhancing the line-to-continuum ratios at high inclinations. 
To construct a scenario where emission 
line equivalent widths are comparable at all inclinations requires 
significant fine-tuning when considering a classical thin disc. 
An alternative, simpler solution is that the emergent continuum is roughly isotropic.

General relativistic effects -- specifically, light bending
and relativistic beaming -- can cause 
the accretion disc SED to become more isotropic \citep[e.g.][]{zhang1997,munozdarias2013}.
To generate GR disc spectra, we use the code \agn\ \citep{hubeny2000,davishubeny2006,davis2007}. 
The output flux at three different wavelengths
as a function of inclination for an \agn\ model with the same disc and BH parameters
as our clumpy wind model is shown in figure~\ref{fig:f2000}.
The effects of GR on an AGN disc are much less extreme 
in the UV portion of the spectrum than the calculations
by \cite{zhang1997} for the X-rays in X-ray binaries.
As a result, the disc emission is still strongly anisotropic.
GR alone therefore cannot explain the line ratio trends in quasars by making the disc
emit more isotropically. This supports the findings of \cite{risaliti2011}, 
who find that EW distributions in quasars are consistent with anisotropic emission 
from optically thick, 
disc-like structures for {\em both} the continuum source and BLR. 
If this is the case, it has a dramatic affect on the intrinsic
BAL fraction inferred from flux-limited samples \citep{goodrich1997,krolikvoit1998}.
We will explore these ideas further in a future study.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/lx.png}
\caption
{
X-ray ($2$~keV) luminosity of the our clumped model (red squares) 
and the H13 model (purple squares), plotted against monochromatic luminosity 
at 2500\AA. The points are labeled according to inclination; angles
$>70^\circ$ correspond to BALs in our scheme (see figure 4).
Also plotted are the samples considered by Saez et al. 2012 on a similar plot; 
The COMBI-7 AGN and the BQS samples Steffen et al. (2006) and the Saez et al. (2012) 
sample of BALQSOs. The dotted line shows the best fit relation for non-BALQSOs 
from Steffen et al. (2006).
}
\label{fig:xray}
\end{figure*} %fullpage

@@
